### 2019-11-02 Rmd/discussion.Rmd
updated ms

## Discussion

### Benefits of our framework

Benefits of our framework

### How does the framework relate to ideas in Nolan and Temple Lang 2010??

### Critiques of our framework

Our measures of students' self-confidence in research ability was merely subjective. In future iterations of our framework, we would like to measure systematic and objective outcomes. One strategy for implementing this is to administer a survey, such as \_\_\_, both before and after the mentored research project. We would use validated survey questions that focused on student beliefs about themselves and their skills.