### 2019-11-21 Rmd/methods.Rmd
updated ms

## Methods

We designed and used a framework for mentored undergraduate data science research 
projects with big data. Our framework draws heavily on ideas in @nolan2010computing. 
Specifically, we incorporate the three key aspects that @nolan2010computing identified:

1. broaden statistical computing to include emerging areas
1. deepen computational reasoning skills
1. combine computational topics with data analysis in the practice of statistics

Additionally, our projects gave students opportunities to develop and to practice skills in reproducible research. Given the growing imperative to document and share code, we feel that this skill set equals in importance the three points above.

4. develop and practice skills in reproducible research

Below, we describe how our framework enabled students to achieve in the four items listed above.

### Broaden statistical computing to include emerging areas

Our framework broadens statistical computing by including the emerging areas of
social media data analysis, sentiment analysis, and topic modeling. 
Both students used Twitter tweets, which we accessed through a Twitter 
streaming API.

Our computational system for acquiring tweets involved several steps. We interacted 
with the API via the R package `twitteR` [@twitteR]. We used the free Twitter streaming 
API that gave us access to approximately one percent stream of all tweets during the 
specified query time period. To ensure that we collected tweets continuously, we used 
the linux tool `crontab` to execute our R script every five minutes. Each execution of 
the R script performed a single streaming API query for five minutes. Twitter's 
streaming API, at the time of our data collection, enforced rate limits on the 
frequency and duration of queries. With the above settings, we continuously collected 
tweets.

### deepen computational reasoning skills

Our framework encourages students to deepen 
computational reasoning skills in several ways. First, they work with a variety of 
internet-based data to answer research questions. In the two example cases, our 
students collected tweets over time and gathered complementary data from other 
resources, including daily closing prices of stock market indexes. This gave students 
opportunities to think creatively about what data to acquire and how to use multiple 
data sources in a single cohesive project.

Second, the students worked with a variety of data structures. 
The Twitter streaming 
API returns tweets as JSON (Javascript Object Notation, http://json.org). Because 
distinct Twitter users may provide 
different pieces of profile information, there is variability in the structure of 
each tweet's JSON. 
Additionally, tweet metadata fields may appear in any order (https://developer.twitter.com/en/docs/tutorials/consuming-streaming-data). 
Students needed to recognize this and to write code that 
accommodated these variations in tweet data structure.

Students wrote code to parse and organize tweet JSONs. They organized their R code into a package, https://github.com/rturn/parseTweetFiles. 
Each tweet's JSON included required fields, and, possibly, some optional fields. 
The required fields are:

**see https://developer.twitter.com/en/docs/tweets/data-dictionary/overview/tweet-object**
**insert a 2-column table of fields here**

The evolving nature of JSON tweet structure 
(https://developer.twitter.com/en/docs/tweets/data-dictionary/guides/tweet-timeline) 
required students to write flexible code that could incorporate newly introduced 
or deprecated metadata.

### combine computational topics with data analysis in the practice of statistics

computational topics: LDA & topic modeling; time series analysis; sentiment analysis.

Both mentored students combined computing with data analysis in the practice of statistics. They used a combination of latent dirichlet allocation topic modeling, sentiment analysis, and time series analysis to reach conclusions about real world data.

Both drew heavily on our collection of tweets. One student examined Standard and Poor's 500 index daily 
closing prices over time. She also analyzed sentiments from each day's stock market-related tweets 
to look for relationships between tweet sentiment and stock market prices.

Our other student focused on developing detection methods for social media events through 
topic modeling of tweets at different time periods. As a proof of principle, he fitted topic 
models to collections of tweets preceding, during, and following the National Football 
League's Super Bowl game. He hypothesized that topics would evolve over time, with football-related tweets appearing during the football game and disappearing soon after conclusion of the game.

data analysis and practice of statistics:

Our mentored research framework begins with brainstorming scientific research ideas 
based on the student's interests. This enables us to craft a project that excites the 
student. With the results of brainstorming sessions, we (mentor and student together) 
formulate the most promising ideas into scientific hypotheses.

For the most appealing 
hypotheses, we encourage the student to translate the scientific question into a 
statistical question that may be addressed with data.

Examples may help to demonstrate our approach to identifying a statistical research 
question. One of our students had interests in acquiring and using social media posts. 
We helped her in brainstorming ideas for research involving social media like Facebook 
and Twitter. Through this brainstorming, we recognized that she had a parallel interest 
in financial markets.

study relationships between social media trends and financial market trends. We helped 
her to brainstorm related research ideas. We then

**justify each step**

**Maybe use/cite Box's science and statistics article from... 1976?**

1. Overview of our framework
  a. student-led, hypothesis-driven research
  b. BH and FB helped with data acquisition
  c. helped with research question formulation
  d. how to describe roles for students v mentors?

Our framework for mentored student research projects involves these components:

1. State the questions and research hypotheses that the two students chose

1. Data description

- how did we/they conceive of projects?

- how did we balance our input with that of the students?
- What data did we use?
- How did we collect data? (which packages, code to sample twitter). How did Jinyu get her stock market data? 
- What is the sampling scheme - for the 1% sample - from the entirety of Twitter? (Cite the webpage that documents the API)