### 2019-11-28 Rmd/outline.Rmd

---
title: "outline"
author: "Frederick J. Boehm"
date: "11/20/2019"
output: bookdown::html_document2
---

## What is the structure of Nolan and Temple Lang 2010??

They give one major section (2, 3, and 4) to each of the three bullet points that I highlight in my methods section.

## What are the important ideas that I need to convey in each section??

**Intro**

Mimic, maybe, the DTL2010 article in organization of the intro.

~~have a subsection on *four* key components (vs DTL's three key components). These are the three of DTL plus reproducible research practices.~~

Maybe also try to write a paragraph about relevant background for each of me and Bret. Note that DTL puts a subsection on backgrounds in their intro.

**Methods**

Here, say more about our contributions - what we did and how we did it. Don't discuss DTL in depth in the Methods - ie, move DTL discussion (that's currently in my methods section) to intro.

Here is the place to discuss our framework and its implementation for two students JX and RT.

**Results**