### 2019-12-02 Rmd/results.Rmd
updates

# Results

We applied the project framework to our mentoring of two students. 
Both engaged in 12 months of mentored research during their senior year of 
undergraduate studies in statistics in 2015 and 2016.

## Student Outcomes

We subjectively assessed student outcomes through conversations in our weekly 
student research meetings. Both students showed increases in confidence and ability to do 
data science research. While the students benefited

Both students secured positions in data science after graduation. 
One student later enrolled in a statistics graduate program, 
while the other pursued employment in health care analytics.

## Mentor Outcomes

## Scholarly outcomes

Our scholarly contributions include the `parseTweetFiles` R package on Github
(https://github.com/rturn/parseTweetFiles) and