### 2019-12-04 Rmd/discussion.Rmd
minor updates to discussion

# Discussion

## Benefits of our framework

1. build self-confidence, proficiency in data analysis and statistics

1. greater interest in quantitative careers and/or grad school in quant fields

1. Working with real data to solve real scientific questions

1. Learn by example interplay between data science (statistics) and science research

Benefits of our framework include enhanced

## How does the framework relate to ideas in Nolan and Temple Lang 2010??

## Critiques of our framework

Our measures of students' self-confidence in research ability was merely subjective. In 
future iterations of our framework, we would like to measure systematic and objective 
outcomes. One strategy for implementing this is to administer a survey, such as \_\_\_, 
both before and after the mentored research project. We would use validated survey 
questions that focused on student beliefs about themselves, their skills, and their 
future careers. **maybe list resources for such survey questions**

## Areas for improvement

1. reproducible research best practices

One shortcoming of our initial framework was the relative lack of emphasis on best practices for 
computational reproducibility. This is one area that we would like to rectify in 
future mentoring activities. The 
University of Wisconsin-Madison has periodically offered a semester course in best practices for 
computationally reproducible research (https://kbroman.org/Tools4RR/). We especially see collaborative 
version control systems, such as Git and 
Github, as essential tools for the modern data scientist.

1. assessment of data science skills

1. assessment of attitudes (pre and post survey??)

1. survey questions might include those from @vance2017eight. I have the survey as a pdf in mendeley. It is part of the supp info for @vance2017eight.