### 2019-12-04 Rmd/results.Rmd

# Outcomes

We applied the project framework to our mentoring of two students. 
Both engaged in 12 months of mentored research during their senior year of 
undergraduate studies in statistics.

## Student Outcomes

We subjectively assessed student outcomes through conversations in our weekly 
student research meetings. Both students showed increases in confidence and ability to do 
data science research.

Both students secured positions in data science after graduation. 
One student later enrolled in a statistics graduate program, 
while the other pursued employment in health care analytics.

## Scholarly outcomes

Our scholarly contributions include the `parseTweetFiles` R package on Github
(https://github.com/rturn/parseTweetFiles) and presentations at conferences such as useR! 2016 and local poster sessions. Additionally, both students prepared end-of-project reports on their research.

We consider below