### 2019-12-06 Rmd/results.Rmd
expanded discussion

# Results

We applied the project framework to our mentoring of two students. 
Both engaged in 12 months of mentored research during their senior year of 
undergraduate studies in statistics. Below, we describe three categories of outcomes:

1. student outcomes
1. mentor outcomes
1. scholarly outcomes

## Student Outcomes

We subjectively assessed student outcomes through conversations in our weekly 
student research meetings. Both students showed increases in confidence and ability to do 
data science research.

Both students secured positions in data science after graduation. 
One student later enrolled in a statistics graduate program, 
while the other pursued employment in health care analytics.

## Mentor outcomes

We (Boehm and Hanlon) grew as mentors during our work with the two students. We successfully
guided junior scientists through a productive, hands-on research experience, and we anticipate
refining the framework in future iterations.

## Scholarly outcomes

Our scholarly contributions include the `parseTweetFiles` R package on Github
(https://github.com/rturn/parseTweetFiles) and presentations at conferences such as useR! 2016 
and local poster sessions. Additionally, both students prepared end-of-project reports on 
their research.