### 2019-12-12 Rmd/results.Rmd
many updates

# Results

We applied the project framework to our mentoring of two students. 
Both engaged in 12 months of research during their senior year of 
undergraduate studies in statistics. Below, we describe three categories of outcomes:

1. student outcomes
1. mentor outcomes
1. scholarly outcomes

## Student Outcomes

We subjectively assessed student outcomes through conversations in our weekly 
student research meetings. Both students showed increases in confidence and ability to do 
data science research.

Both students secured positions in data science after graduation. 
One student enrolled in a statistics graduate program, 
while the other pursued employment in health care analytics.

Students benefited from our emphasis on the four central concepts, 
three from @nolan2010computing plus reproducible research skills. 
The research projects successfully drew on emerging areas of 
statistical computing, namely text analysis. They combined 
computational topics, including topic modeling and time series 
methods, with data analysis in the practice of statistics. Although 
we didn't formally measure them, our informal assessment indicates 
that students' computational reasoning skills increased over the 
duration of our projects. Students used a variety of computing tools 
and methods to arrive at a practical solution to a selected task. 
They became more skilled in computing with R and shell scripts and 
more fluid in their verbal explanations during our regular meetings [@r].

Our framework's emphasis on reproducible research skills is 
evidenced by the students' R package, `parseTweetFiles`, which is 
both version controlled with `git` and shared via Github.com.

## Mentor outcomes

We (Boehm and Hanlon) grew as mentors during our work with the two students. We successfully
guided junior scientists through a productive, hands-on research experience, and we anticipate
refining the framework in future iterations.

## Scholarly outcomes

Our scholarly contributions include the `parseTweetFiles` R package on Github
(https://github.com/rturn/parseTweetFiles) and presentations at conferences such as useR! 2016 (R users' conference)
and local poster sessions. Additionally, both students prepared end-of-project reports on 
their research.