### 2019-12-14 Rmd/cover-letter.Rmd
added submission materials.

---
title: ""
author: ""
date: "12/14/2019"
output: github_document
---

Dear editors:

We're writing to submit a manuscript for the Journal of Statistics Education's special issue on "Computing in the 
Statistics Curriculum". In it, we describe our mentoring framework for undergraduate research projects in data science.
We relate the framework to many ideas from Nolan & Temple Lang (2010) and offer minor amendments to Nolan & Temple Lang's 
arguments.

We developed the framework while guiding two undergraduate students through their senior projects in statistics.

Our framework integrates ideas from Nolan and Temple Lang (2010) with literature on mentoring and curriculum design.

We thank you for your careful consideration of our manuscript.

Sincerely,

Frederick Boehm and Bret Hanlon