### 2020-05-24 Rmd/abstract.Rmd
updated ms

# Abstract

We describe methods for collecting tweets from Twitter. With these methods, we 
curate a large collection of tweets.
We present methods for social media event detection through text analysis of tweets. 
We describe our use of tweets in mentored undergraduate research. 
We also offer other directions for text analysis of tweets.