### 2020-05-29 Rmd/tweets.Rmd

---
title: "Tweets"
author: "Frederick J. Boehm"
date: "4/9/2020"
output: html_document
---

## R code to query twitter streaming api

Install `rtweet` package from CRAN.

```{r, eval = FALSE}
install.packages("rtweet")
install.packages("lubridate")
```

## Parse json file

```{r}
library(magrittr)
source("recover_tweets.R")
```

```{r, stops}
# create a twitter-specific stop word list
remove_reg <- "&amp;|&lt;|&gt;"
twitter_stop <- tibble::tibble(word = c("https", "http", "t.co", "amp"))
# get a list of english language words and names
library(qdapDictionaries) # contains object GradyAugmented
library(tidytext) # get stop_words
```

```{r}
date <- c("2020-05-24", "2020-05-25", "2020-05-26")
tweet_json_file <- c("../data/2020-05-24 12:00:02-tweets.json",
  "../data/2020-05-25 12:00:02-tweets.json", 
  "../data/2020-05-26 12:00:02-tweets.json"
  )
out <- purrr::pmap_dfc(tibble::tibble(date, tweet_json_file),
                       .f = function(date, tweet_json_file){
                         knitr::knit_expand(text = readLines("tweets-one.Rmd"))
                       })
```

## May 24, 2020

`r knitr::knit(text = out[[1]])`

## May 25, 2020

`r knitr::knit(text = out[[2]])`

## May 26, 2020

`r knitr::knit(text = out[[3]])`

## Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```