### 2020-05-29 Rmd/abstract.Rmd
revised abstract. added section numbers

# Abstract

We present a freely available, large, rich data set from Twitter and describe how we collected 
it. We then describe mentored student research that uses tweets to address novel research
questions. We use backward design principles to develop learning objectives
for student researchers. We conduct formative and summative assessments of student learning.
To illustrate the value of our data set, we 
present details of one student project. We conclude by discussing future research 
directions.