### 2020-06-01 Rmd/abstract.Rmd
fixed abstract

# Abstract

We present a framework for undergraduate research projects with Twitter data. 
Drawing on backward design principles, we share our learning objectives and
summative assessments. We draw on experiences mentoring two students during 
an academic year. To illustrate the value of Twitter as a data source, 
we detail methods for collecting and analyzing tweets. We conclude by emphasizing
how Twitter text analysis projects enable students to formulate 
original research questions, collect and analyze the needed data, and communicate
findings and their implications.