### 2020-08-14 Rmd/abstract.Rmd

# Abstract

We draw on our experiences with mentoring two students to develop a framework for undergraduate research projects with Twitter data.
Leveraging backward design principles, we share our learning objectives and
rubric for summative assessments. To illustrate the value of Twitter as a data source, 
we detail methods for collecting and analyzing tweets. We conclude by emphasizing
how Twitter text analysis projects enable students to formulate 
original research questions, collect and analyze data, and communicate
findings and their implications.