### 2020-10-07 Rmd/jse-special-issue-2020-october.tex
fixed yaml to keep tex file

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={What is happening on Twitter? A framework for student research projects with tweets},
  pdfauthor={Frederick J. Boehm and Bret M. Hanlon},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
%\usepackage{lineno}
%\linenumbers
\usepackage{setspace}
\doublespacing
\usepackage{caption}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
%\usepackage[utf8x]{inputenc}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{What is happening on Twitter? A framework for student research projects with tweets}
\author{Frederick J. Boehm and Bret M. Hanlon}
\date{}

\begin{document}
\maketitle

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

We draw on our experiences with mentoring two students to develop a framework for undergraduate research projects with Twitter data.
Leveraging backward design principles, we share our learning objectives and
rubric for summative assessments. To illustrate the value of Twitter as a data source,
we detail methods for collecting and analyzing tweets. We conclude by emphasizing
how Twitter text analysis projects enable students to formulate
original research questions, collect and analyze data, and communicate
findings and their implications.

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Twitter has profoundly changed how we communicate.
In only 280 characters, users instantly contribute to public conversations on politics,
current events, sports, media, and many other topics.
Recent development of accessible statistical methods for large-scale text
analysis now enable instructors to use tweets as contemporary pedagogical
tools in guiding undergraduate research projects.
We guided two statistics students in their senior research projects.
Both students used tweets to address novel research questions. We
share products of their research in supplementary files. Because
their data are no longer available, we present as a case study one
analysis with tweets from May 2020. We share our data and computer
code to encourage others to undertake
tweet text analysis research. We also describe methods for creating a
collection of tweets.

Some social media data, including tweets from Twitter, is
available through website application product interfaces (APIs).
By way of a streaming API, Twitter shares a sample of approximately
one percent of all
tweets during an API query time period (``Sampled stream'' 2019). Any Twitter
user can freely
access this one percent sample, whereas access to a larger selection is available to
researchers for a fee.

Through our work with tweets, we demonstrate
that Twitter data is a rich source of new data science research questions.
Box (1976) described a
positive feedback loop for the interplay of discipline-specific research
and quantitative methods research. The two components, ``science'' and ``statistics'' in the
language of Box (1976), iteratively fuel research questions in each other.
A new statistical
method enables new discipline-specific questions to be addressed, while a new
scientific question motivates new data science methods. We describe below two
data science research questions that mentored students addressed with tweets.

Studies of Twitter conversations have yielded valuable insights into
modern culture. Using large collections of tweets, scholars have
investigated diverse research questions, including
the inference of relationships and social networks among Twitter users (Lin et al. 2011);
authorship of specific tweets when multiple persons share a single account (Robinson 2016); and
rhetoric in recruiting political supporters (Pelled et al. 2018; Wells et al. 2016).
Recognizing the potential utility of tweets for data science research and teaching,
we created a collection of tweets over time by repeated querying of the Twitter streaming API.
We envisioned this collection as a rich resource for data science research projects. This vision
grew into two mentored undergraduate student research projects in the 2015-2016 academic year.

In line with recent calls for students to work with real data (Carver et al. 2016; Nolan and Temple Lang 2010), our
collection of tweets has served as a valuable resource in our mentoring
of undergraduate data science research.
Working with real data allows students to develop proficiency not
only in statistical analysis,
but also in related data science skills, including data
transfer from online sources, data storage, using data
from multiple file formats, and communicating findings.
Collaboratively asking and addressing novel questions with
our collection of tweets gave mentored students
opportunities to develop competency in all of these areas.

While our tweet collection enables us to address many possible research questions, the
dynamic content of tweets over time particularly piqued our interest.
We hypothesized that high-profile social media events would generate a high volume of
tweets, and that we would detect social media events through changes in tweet
topic content over time. We discuss in detail below one approach to studying this
question. In the sections that follow, we detail our backward design-inspired approach to
writing learning objectives, preliminary research mentoring considerations,
data science methods for collecting and analyzing tweets, analysis results,
and ideas on assessment and advanced topics.

\hypertarget{structure-of-mentored-research}{%
\section{Structure of mentored research}\label{structure-of-mentored-research}}

\hypertarget{backward-design}{%
\subsection{Backward design}\label{backward-design}}

Backward design principles guided our planning and informed the writing of
learning objectives (Wiggins and McTighe 2005). Following Wiggins and McTighe (2005),
we began by listing
what students, at the end of their thesis research, should be able to do,
understand, and know. We then classified each of these items into
one of three categories: enduring understanding, important to know and do,
and worth being familiar with (Wiggins and McTighe 2005) (Table \ref{tab:circle-table}).
While other researchers may categorize these skills differently,
our assignments reflect our projects' priorities.
Nearly all of the skills in Table \ref{tab:circle-table} are
transferable. They apply not merely to thesis projects, but also to
data science research in general.

\begin{table}

\caption{\label{tab:circle-table}Classifying project skills}
\centering
\begin{tabular}[t]{|>{}l|>{}l|}
\hline
Skill & Category\\
\hline
Structure research project files as R package & Worth being familiar with\\
\hline
Communicate results in speaking and in writing & Enduring understanding\\
\hline
Formulate a research question & Enduring understanding\\
\hline
Develop data science strategies to address research question & Enduring understanding\\
\hline
Use Github to share code and documentation & Important to know and do\\
\hline
Use git for version control & Important to know and do\\
\hline
Use text analysis tools to analyze tweets & Enduring understanding\\
\hline
Use cluster computing as needed & Worth being familiar with\\
\hline
Use data visualization to clarify and inform quantitative analyses & Important to know and do\\
\hline
Translate analysis results into scientific conclusions & Enduring understanding\\
\hline
Incorporate supplementary data sources into analysis & Important to know and do\\
\hline
Acquire data from internet sources & Important to know and do\\
\hline
Describe assumptions and limitations of statistical analyses & Enduring understanding\\
\hline
\end{tabular}
\end{table}

In particular, our project skills reflect elements of ``data acumen'', as defined
in a report from the U.S.A. National Academies of Sciences, Engineering, and Medicine (NASEM) (National Academies of Sciences et al. 2018). For example, our skills ``Acquire data from internet sources'' and
``Develop data science strategies to address a research question'' implicitly require
a trainee, in the language of the NASEM report (National Academies of Sciences et al. 2018), to ``ingest, clean, and then
wrangle data into reliable and useful forms'' and to ``combine many existing programs or codes
into a workflow that will accomplish some important task''. Additionally, we tailored our list
of project skills with the assumption that students would work in R. R use is not required for such projects, but it is a
convenience for many of our students.

\hypertarget{learning-objectives}{%
\subsection{Learning objectives}\label{learning-objectives}}

We translated our prioritized list of skills that
students should be able to do, understand, and know into learning objectives (Table \ref{tab:circle-table}).
We phrased learning objectives in a manner that enabled their subsequent assessment
(Table \ref{tab:summative}) with formative and summative strategies. These were
our four learning objectives:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Write R code to perform text analysis of large volumes of tweets (R Core Team 2019).
\item
  Communicate results in a written report and poster presentation.
\item
  Translate statistical findings into scientific conclusions.
\item
  Develop data science strategies to address a scientific research question.
\end{enumerate}

Having decided on four learning objectives for students, we next established
mentoring relationships with each student.

\hypertarget{preliminary-research-mentoring-considerations}{%
\subsection{Preliminary research mentoring considerations}\label{preliminary-research-mentoring-considerations}}

We developed research goals with students in a series of brainstorming sessions and discussions.
As trainees began their senior
research projects, we spoke in detail about both their research interests and goals and their
experience with data analysis software. When possible, we encouraged them to incorporate their
existing academic interests into their senior research projects.

In our statistics department, most students learn elementary R computing skills through class
assignments. Some students, by concentrating in computer science, learn other data analysis software
packages, such as Python.
Those who do undergraduate statistics research often learn advanced topics in R
computing, such as R package assembly, documentation, and testing. Many develop
expertise in linux computing and cluster computing, too.

One of our two students had extensive experience in statistical computing. In
addition to R computing skills, she also worked in Python and excelled in shell scripting.
She first learned Python in computer science courses. Our second student had
extensive experience with R from his statistics courses. His background
enabled him to write an R package as part of his senior project. To encourage
further development of R computing skills in our two students, we guided them
towards the free, online books ``R for Data Science'' (Wickham and Grolemund 2016) and ``Advanced R'' (Wickham 2019).

\hypertarget{student-research-interests-and-goals}{%
\subsection{Student research interests and goals}\label{student-research-interests-and-goals}}

Our two students had diverse interests, and, initially, they had little experience in
articulating research goals. We engaged each in a brainstorming session to
clarify their interests and encourage them to think critically about research goals under
the time constraints of their academic schedules.
We briefly describe the two student projects to give readers a better sense
of research possibilities with tweets.

Our first student examined relationships over time between
stock market index prices and tweet sentiment. For each day in her 12-month study period,
she identified stock market-related tweets with a key word search. With the complete
texts of stock market-related tweets for each day, she calculated a daily sentiment score and plotted it over time.
Her sentiment score reflected presence of emotion-associated terms
(\emph{eg}., ``happy'', ``sad'', ``mad'', ``scared'')
in tweet texts. Days with more net positive emotion words in the collected tweets received a higher
(positive) daily sentiment score, while days with more net negative words received a negative daily
sentiment score.
For her final project, she presented plots over time of her daily sentiment scores and daily closing
prices of the Standard and Poor's 500 index. She also explored time series analysis methods to quantify
relationships between index prices and sentiment scores.

Our second student developed social media event detection methods with topic
models. He hypothesized that tweet content changes over time, and that we
might detect these changes by comparing inferred tweet topics from distinct
time periods. To validate his hypothesis, he examined tweet content before, during, and after
the National Football League's Super Bowl game in 2015. He reasoned that
because the Super Bowl is widely discussed on Twitter, we might detect
Super Bowl-related topics from tweets sent during the game, but that the football-related
topics would be short-lived in the continuous Twitter stream. We discovered
evidence to support his strategy, and we ultimately presented our findings
at international and local research meetings. We share below a case study based
on this approach.

\hypertarget{time-period}{%
\subsection{Time period}\label{time-period}}

Our two statistics students conducted their research projects during the 2015-2016 academic year.
We recommend a full academic year for projects of this magnitude, although a summer or one-semester
project is possible. Our students presented their findings at the statistics department's
undergraduate poster session near the end of the 2015-2016 academic year (Supplementary files). We
present below reproducible R code for analyzing data from May 2020. While these
are not the same data that our students analyzed in 2015, the methods and code
are very similar to that of our second student's project.

\hypertarget{case-study-methods}{%
\section{Case study methods}\label{case-study-methods}}

To illustrate the value of Twitter data and to encourage readers to envision other uses for tweets, we present below a
reproducible case study. It is essentially a reproduction of our second student's project, but at a
distinct time period. In it, we aim to detect a social media
event by examining tweet topic content over time. We use Latent Dirichlet
Allocation (LDA) models (Blei et al. 2003) to infer topics on three consecutive days centered
on Memorial Day 2020. We chose this example case study,
instead of the student projects, because of limited data
availability for the student projects. Despite this, the case study
illustrates the strategy and methods for one student
project. Below, we discuss case study design, tweet collection, and tweet structure,
before turning to quantitative methods for the case study.

\hypertarget{case-study-design}{%
\subsection{Case study design}\label{case-study-design}}

We sought to validate our hypothesis that we could detect a social
media event by examining tweet topic content at distinct time periods.
As a proof of principle of our event detection strategy,
we analyzed tweets before, during, and
after Memorial Day (May 25, 2020).
We fitted LDA models for each of three
distinct five-minute periods. The first period began at noon Eastern time
on May 24, 2020. Subsequent time periods started
24 and 48 hours later.
We defined each time period to be a single collection, or corpus, of
tweets.

\hypertarget{collecting-tweets-over-time}{%
\subsection{Collecting tweets over time}\label{collecting-tweets-over-time}}

We include here instructions for creating a tweet collection. First, we created a
new account on Twitter. With these user credentials, we used the
R package \texttt{rtweet} to query the API (Kearney 2019). We used the linux \texttt{crontab} software
to repeatedly execute R code to submit API queries. Each
query lasted five minutes and produced a text file of approximately 130 MB. We timed the API queries so
that there was no time lag between queries. We stored tweets resulting from API queries in their native JSON format.

Setting up the query task with \texttt{crontab} is straightforward. On our computer, with
Ubuntu 20.04 linux operating system, we opened a terminal and typed \texttt{crontab\ -e}. This
opened a text file containing user-specified tasks. We added the following line to the
bottom of the file before saving and closing the text file.

\begin{Shaded}
\begin{Highlighting}[]
\ExtensionTok{*/5}\NormalTok{ * * * * R -e }\StringTok{'rtweet::stream_tweets(timeout = (60 * 5), }
\StringTok{parse = FALSE, file_name = paste0("~/work/mentoring/mentoring-framework/data/",}
\StringTok{lubridate::now(), "-tweets"))'}
\end{Highlighting}
\end{Shaded}

Readers may need to slightly amend the above line to conform to requirements of
their operating system's software. Readers who use Mac OS may proceed as we did,
while those with Windows operating systems may consider using the R package
``taskscheduleR'' to schedule API queries via the Windows task scheduler (Wijffels and Belmans 2018).

\hypertarget{querying-twitter-api-to-get-complete-tweets}{%
\subsection{Querying Twitter API to get complete tweets}\label{querying-twitter-api-to-get-complete-tweets}}

Twitter API use agreements forbid users from sharing complete API query results.
However, Twitter permits users to share tweet identification numbers. With a tweet
identification number, a user may query a Twitter API to obtain complete tweet data. In our
experience, this
process is incomplete; that is, many tweet identification numbers submitted to the Twitter API return
no data. Additionally, some tweet identification numbers return data on the first query, but don't return data on subsequent queries. This complicates our goal of making all analyses computationally
reproducible and motivates our decision to share the tweet IDs of those tweets that
we actually analyzed (Supplementary files). Should a reader wish to reproduce
our analysis, we anticipate that she will get complete tweet data
for all or most of these tweet identification numbers from the API. We provide
R code for this task in the supplementary files.

\hypertarget{tweet-structure}{%
\subsection{Tweet structure}\label{tweet-structure}}

Tweets are available from the Twitter API as Javascript Object Notation (JSON) objects (``Introducing JSON'' 2020).
Every tweet consists
of multiple key-value pairs. The number of fields per
tweet depends on user settings, retweet status, and other factors (``Introduction to Tweet JSON'' 2020).
The 31 tweet key-value pairs belong to
12 distinct classes (Supplementary files). The
classes are either vectors - numeric, logical, or character - or arrays assembled
from the vector classes.

Below is an example of JSON for one tweet.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{\{}
  \StringTok{"created_at"}\NormalTok{: }\StringTok{"Thu Apr 06 15:24:15 +0000 2017"}\NormalTok{,}
  \StringTok{"id_str"}\NormalTok{: }\StringTok{"850006245121695744"}\NormalTok{,}
  \StringTok{"text"}\NormalTok{: }\StringTok{"1\textbackslash{}/ Today we\textbackslash{}u2019re sharing our vision for the future of the Twitter API platform!"}\NormalTok{,}
  \StringTok{"user"}\NormalTok{: }\KeywordTok{\{}
    \StringTok{"id"}\NormalTok{: }\ExtensionTok{2244994945}\NormalTok{,}
    \StringTok{"name"}\NormalTok{: }\StringTok{"Twitter Dev"}\NormalTok{,}
    \StringTok{"screen_name"}\NormalTok{: }\StringTok{"TwitterDev"}\NormalTok{,}
    \StringTok{"location"}\NormalTok{: }\StringTok{"Internet"}\NormalTok{,}
    \StringTok{"url"}\NormalTok{: }\StringTok{"https:\textbackslash{}/\textbackslash{}/dev.twitter.com\textbackslash{}/"}\NormalTok{,}
    \StringTok{"description"}\NormalTok{: }\StringTok{"Your official source for Twitter Platform news, updates & events. }
\StringTok{    Need technical help? Visit https:\textbackslash{}/\textbackslash{}/twittercommunity.com\textbackslash{}/ \textbackslash{}u2328\textbackslash{}ufe0f }
\StringTok{    #TapIntoTwitter"}
  \KeywordTok{\}}\NormalTok{,}
  \StringTok{"place"}\NormalTok{: }\KeywordTok{\{}   
  \KeywordTok{\}}\NormalTok{,}
  \StringTok{"entities"}\NormalTok{: }\KeywordTok{\{}
    \StringTok{"hashtags"}\NormalTok{:}\BuiltInTok{ [}      
\NormalTok{    ],}
    \StringTok{"urls"}\NormalTok{: [}
\NormalTok{      \{}
        \StringTok{"url"}\NormalTok{: }\StringTok{"https:\textbackslash{}/\textbackslash{}/t.co\textbackslash{}/XweGngmxlP"}\NormalTok{,}
        \StringTok{"unwound"}\NormalTok{: \{}
          \StringTok{"url"}\NormalTok{: }\StringTok{"https:\textbackslash{}/\textbackslash{}/cards.twitter.com\textbackslash{}/cards\textbackslash{}/18ce53wgo4h\textbackslash{}/3xo1c"}\NormalTok{,}
          \StringTok{"title"}\NormalTok{: }\StringTok{"Building the Future of the Twitter API Platform"}
\NormalTok{        \}}
\NormalTok{      \}}
\NormalTok{    ],}
    \StringTok{"user_mentions"}\NormalTok{: [     }
\NormalTok{    ]}
\NormalTok{  \}}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

Our analyses use three fields from each tweet: date (``created\_at''), tweet identifier
(``id\_str''), and tweet text (``text''). The ``created\_at'' field is a character string containing
the date and time of the tweet. Every tweet has a unique identifier, the ``id\_str'' value. The
``text'' field contains the unicode representation of the message. After creating a text file
with tweet JSON, our next step involved reading and parsing tweets with the R
packages \texttt{rtweet} (Kearney 2019) and \texttt{tidytext} (Silge and Robinson 2016).

\hypertarget{parsing-tweet-text}{%
\subsection{Parsing tweet text}\label{parsing-tweet-text}}

The next task is to wrangle the tweet JSON into a data structure suitable for LDA modeling.
We used functions from the \texttt{rtweet} R package to parse tweet JSON into a data frame.
We then divided tweet text into words with functions from the \texttt{tidytext} R package. We
discarded commonly used ``stop words'' and emojis.

LDA model fitting requires that the corpus be organized as a
document by term matrix. In a document by term matrix, each row corresponds to a single document (a single
tweet), and each column is a single term (or word). Each cell contains a
count (the number of
occurrences of a term in the specified document). We created a document by
term matrix with the R function \texttt{cast\_dtm} from the \texttt{tidytext} package.

\hypertarget{latent-dirichlet-allocation}{%
\subsection{Latent Dirichlet Allocation}\label{latent-dirichlet-allocation}}

LDA is a statistical method for inferring latent (unobservable) topics (or themes)
from a large corpus (or collection) of documents (Blei et al. 2003).
We pretend that there's an imaginary process for creating documents in the corpus. For each document,
we choose a discrete distribution over topics. For example, some tweets from Memorial Day may
refer to the holiday. This may constitute one topic in the corpus.
Having chosen a distribution over topics, we then select document words by first
drawing a topic from the distribution over topics, then drawing a word from the
chosen topic.

In mathematical notation, we write the generative process assumed by LDA (Blei et al. 2003):

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Choose \(N \sim \text{Poisson}(\xi)\)\\
\item
  Choose \(\theta \sim \text{Dirichlet}(\alpha)\)\\
\item
  For each word, \(w_n\) with \(n = 1, ..., N\),
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\alph{enumi}.}
\tightlist
\item
  Choose a topic \(z_n \sim \text{Multinomial}(\theta)\)\\
\item
  Choose a word \(w_n\) from \(p(w_n | z_n, \beta)\), a multinomial probability
\end{enumerate}

\(\beta\) refers to the \(k\) by \(V\) matrix of topic-specific word probabilities, where \(k\)
is the number of topics and \(V\) is the size of the vocabulary, \emph{i.e.}, the number of unique words in the corpus.

The goal for LDA is to infer both the distribution over topics and
the topics (Blei et al. 2003). A topic, in this setting, is a distribution over
the vocabulary (the collection of all words in a corpus).
Inference for latent Dirichlet allocation models is performed by either
sampling from the posterior distribution or through variational methods. Researchers
have devised a variety of Gibbs sampling techniques for these models (Porteous et al. 2008).
Variational methods, while using approximations to the posterior distribution, offer the
advantage of computational speed (Blei et al. 2017). We used variational methods below in our case study.

\hypertarget{case-study-results}{%
\section{Case study results}\label{case-study-results}}

We identified the top ten most probable terms for each of ten topics in our models
(Figures \ref{fig:may24}, \ref{fig:may25}, \ref{fig:may26}).
We plotted the within-topic word probabilities as bar graphs. We see that topic-specific
word probabilities seldom exceed 0.05. We also note that some words are heavily weighted in
multiple topics. This observation complicates semantic topic interpretation. We
also caution that the results display expletives (that appeared on Twitter) and may
not be ``safe for work''.

\begin{figure}
\includegraphics[width=29.17in]{../results/beta-2020-05-24} \caption{Top terms for LDA model from May 24, 2020}\label{fig:may24}
\end{figure}

\begin{figure}
\includegraphics[width=29.17in]{../results/beta-2020-05-25} \caption{Top terms for LDA model from May 25, 2020 (Memorial Day)}\label{fig:may25}
\end{figure}

\begin{figure}
\includegraphics[width=29.17in]{../results/beta-2020-05-26} \caption{Top terms for LDA model from May 26, 2020}\label{fig:may26}
\end{figure}

Assigning meaning to topics is an active research area (Chang et al. 2009). Since
our interest is in the transient appearance of a new topic, we don't attempt to
assign meaning to every topic in our models. Instead, we anticipate that discussions on Twitter are a mixture of topics that endure over weeks or months and subjects that appear and disappear quickly. We see that topic 7 from May 25 has several
words that suggest Memorial Day: memorial, remember, honor, country. A similar topic is not
seen on May 24 or May 26. Some topics persist, with distinct word probabilities,
across the three days. For example, we see that President Trump features prominently in all
three models' results. On May 26, topic 10 reflects discussion of the Amy
Cooper Central Park incident (\url{https://www.nytimes.com/2020/05/26/nyregion/amy-cooper-dog-central-park.html}).

The murder of George Floyd occurred on May 25, 2020. Our last
examined time period, from 12:00 pm to 12:05 pm (Eastern USA time zone) on May 26, occurred after Floyd's murder,
yet we didn't detect this event in our ten-topic LDA model. Several considerations
may account for this. While outrage at the murder eventually spread worldwide, there may have been few
Floyd-related tweets during our collection time on May 26, less than 24 hours after the murder and video release.
Had we extended our analysis to May 27 and beyond, we may have identified George Floyd-related topics.

\hypertarget{assessment-of-learning-exploring-more-advanced-topics-and-concluding-remarks}{%
\section{Assessment of learning, exploring more advanced topics, and concluding remarks}\label{assessment-of-learning-exploring-more-advanced-topics-and-concluding-remarks}}

\hypertarget{assessment-of-learning}{%
\subsection{Assessment of learning}\label{assessment-of-learning}}

We examined student learning with both formative and summative assessments.
We conducted formative assessments through weekly discussions with students.
In these discussions, we developed action items to advance research progress and overcome
challenges. We summatively assessed student achievement at the end of the academic year.
Both students wrote a thesis and presented a poster to our statistics department.
We asked questions at the poster session to probe student understanding and critically
evaluated the theses.

\begin{table}

\caption{\label{tab:summative}Rubric for summative assessment of learning objectives.}
\centering
\fontsize{6}{8}\selectfont
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{10em}|>{\raggedright\arraybackslash}p{10em}|>{\raggedright\arraybackslash}p{10em}|>{\raggedright\arraybackslash}p{10em}|>{\raggedright\arraybackslash}p{10em}}
\hline
Learning objective & Assessment item & 2 points & 1 point & 0 points\\
\hline
Write R code to perform text analysis of large volumes of tweets. & R code performs intended analyses & Code contains few or no bugs & Code contains one or more errors & Code contains many errors\\
\hline
Write R code to perform text analysis of large volumes of tweets. & Uses literate programming tools, such as Sweave or knitr & Report is written using literate programming tools. It compiles easily when run by instructor. Time-consuming calculations are cached. & Report is written using literate programming tools, but compilation takes too long or fails. & Report is not written with literate programming tools.\\
\hline
Write R code to perform text analysis of large volumes of tweets. & Uses git for version control & Log reveals regular commits with informative commit messages & Log reveals intermittent commits and uninformative commit messages & Doesn't use git.\\
\hline
Write R code to perform text analysis of large volumes of tweets. & Shares code and data via Github & Instructor easily clones repository from Github. Contains share-able data and instructions for getting other data to reproduce analysis. & One or more needed files is missing from repository. & Doesn't use Github.\\
\hline
Communicate results in a written report and poster presentation. & Organizes poster to highlight main points & When prompted, can describe main points in less than one minute. & Less fluid presentation with periods of silence or confusion. & Disorganized presentation.\\
\hline
Communicate results in a written report and poster presentation. & Accurately presents study and findings during poster session & Fluently describes background, study goals, study design, approach, data, findings, and conclusions & At least one section is incomplete or is verbal explanation is incomplete. & At least one section is missing.\\
\hline
Communicate results in a written report and poster presentation. & Report structure mirrors a research manuscript & Contains abstract, introduction, methods, results, and discussion & At least one section is incomplete. & At least one section is missing.\\
\hline
Translate statistical findings into scientific conclusions. & Places statistical results in their scientific context & Demonstrates understanding of scientific context and integrates findings into it. & Incomplete scientific understanding or incomplete integration of findings. & Major gaps in scientific understanding or integration of findings.\\
\hline
Translate statistical findings into scientific conclusions. & Accurately portrays study limitations & Accurately describes, in writing and in speaking, limitations of the study & Incomplete or partially inaccurate description of limitations & Doesn't describe limitations.\\
\hline
Translate statistical findings into scientific conclusions. & Demonstrates familiarity with relevant literature & Fluent in both relevant data science literature and scientific literature. & Incomplete knowledge and understanding of relevant literature & Major gaps in knowledge and understanding\\
\hline
Develop data science strategies to address a scientific research question. & Presents an original research question & Presents, in writing and in speaking, a novel research question. Explains why it's novel, too. & Partially lacking in elements of question's background or novelty. & Doesn't present an original question.\\
\hline
Develop data science strategies to address a scientific research question. & Effectively uses data visualizations & Visualizations highlight main points of report. & Incomplete or omitted visualizations. & Doesn't use visualizations.\\
\hline
Develop data science strategies to address a scientific research question. & Presents accurate scientific conclusions & Effectively translates analysis results into their scientific context. & Minor inaccuracy in translation of findings into scientific context. & Major errors in translation of results.\\
\hline
\end{tabular}
\end{table}

With future students, we will use a written rubric to evaluate
theses (Table \ref{tab:summative}). We'll
share the rubric with our students at the start of the academic year. With only minor
modifications, the rubric may be suitable for projects that don't use tweets.

\hypertarget{exploring-more-advanced-topics}{%
\subsection{Exploring more advanced topics}\label{exploring-more-advanced-topics}}

Twitter data over time inspires a variety of research projects. Supplementing
tweets with public data from other sources multiplies the possibilities. For example,
one of our two students supplemented tweets with daily stock market index prices.
She studied sentiment of finance-related tweets and daily stock market index
closing prices (Supplementary files).

LDA modeling and related methods are a major research area in the
quantitative social sciences. Advanced students with interest in statistical computing might
compare inferential methods for topic models. Those with interests in event detection
and time series analysis could build on the findings of our student by explicitly accounting
for topic evolution with dynamic topic models (Blei and Lafferty 2006).

\hypertarget{concluding-remarks}{%
\subsection{Concluding remarks}\label{concluding-remarks}}

Our mentoring in data science aligns with others' calls to reconsider
the role of computing in statistics
and data science (Carver et al. 2016; Nolan and Temple Lang 2010). Hicks and Irizarry (2018) argue for
incorporating three concepts into data science training:
computing, connecting and creating.
They use the terms ``connecting'' and ``creating'' to describe the processes of applying
quantitative methods to real data and research questions and of formulating research
questions, respectively. Our tweet analysis projects offer students opportunities in all
three skills sets. Our students first formulated research questions, then collected and
analyzed data to address the questions. Throughout the projects, students
drew heavily on computing, both to acquire data and to analyze it.

Tweet analysis gives students practical experience in the data science process of
formulating a research question, gathering data to address it, summarizing the
data, visualizing results, and communicating findings. Tweets over time are a rich, large,
authentic data set that offers many opportunities.
We provided instructions to enable readers to establish their own tweet collections.
We also presented details for one analysis strategy.
By considering first student research interests and integrating them with our
senior thesis learning objectives, we successfully guided two undergraduate
researchers in data science research with tweets.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

The authors thank Betsy Colby Davie and Rick Nordheim for helpful discussions and
feedback on preliminary versions of the manuscript. We thank the special issue
editors and anonymous reviewers for their constructive comments and suggestions. Finally, this work wouldn't have been possible without the keen and enthusiastic students, Jinyu Xia and Robert Turner.

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-blei2017variational}{}%
Blei, D. M., Kucukelbir, A., and McAuliffe, J. D. (2017), ``Variational inference: A review for statisticians,'' \emph{Journal of the American Statistical Association}, Taylor \& Francis, 112, 859--877.

\leavevmode\hypertarget{ref-blei2006dynamic}{}%
Blei, D. M., and Lafferty, J. D. (2006), ``Dynamic topic models,'' in \emph{Proceedings of the 23rd International Conference on Machine Learning}, pp. 113--120.

\leavevmode\hypertarget{ref-blei2003latent}{}%
Blei, D. M., Ng, A. Y., and Jordan, M. I. (2003), ``Latent Dirichlet Allocation,'' \emph{Journal of Machine Learning Research}, 3, 993--1022.

\leavevmode\hypertarget{ref-box1976science}{}%
Box, G. E. (1976), ``Science and statistics,'' \emph{Journal of the American Statistical Association}, Taylor \& Francis, 71, 791--799.

\leavevmode\hypertarget{ref-carver2016guidelines}{}%
Carver, R., Everson, M., Gabrosek, J., Horton, N., Lock, R., Mocko, M., Rossman, A., Roswell, G. H., Velleman, P., Witmer, J., and others (2016), ``Guidelines for assessment and instruction in statistics education (GAISE) college report 2016,'' AMSTAT.

\leavevmode\hypertarget{ref-chang2009reading}{}%
Chang, J., Gerrish, S., Wang, C., Boyd-Graber, J. L., and Blei, D. M. (2009), ``Reading tea leaves: How humans interpret topic models,'' in \emph{Advances in Neural Information Processing Systems}, pp. 288--296.

\leavevmode\hypertarget{ref-hicks2018guide}{}%
Hicks, S. C., and Irizarry, R. A. (2018), ``A guide to teaching data science,'' \emph{The American Statistician}, Taylor \& Francis, 72, 382--391.

\leavevmode\hypertarget{ref-json}{}%
``Introducing JSON'' (2020), \url{https://www.json.org/json-en.html}.

\leavevmode\hypertarget{ref-tweet_json}{}%
``Introduction to Tweet JSON'' (2020), \url{https://developer.twitter.com/en/docs/tweets/data-dictionary/overview/intro-to-tweet-json}.

\leavevmode\hypertarget{ref-rtweet-package}{}%
Kearney, M. W. (2019), ``rtweet: Collecting and analyzing Twitter data,'' \emph{Journal of Open Source Software}, 4, 1829. \url{https://doi.org/10.21105/joss.01829}.

\leavevmode\hypertarget{ref-lin2011joint}{}%
Lin, C. X., Mei, Q., Han, J., Jiang, Y., and Danilevsky, M. (2011), ``The joint inference of topic diffusion and evolution in social communities,'' in \emph{2011 IEEE 11th International Conference on Data Mining}, IEEE, pp. 378--387.

\leavevmode\hypertarget{ref-national2018data}{}%
National Academies of Sciences, Engineering, Medicine, and others (2018), \emph{Data science for undergraduates: Opportunities and options}, National Academies Press.

\leavevmode\hypertarget{ref-nolan2010computing}{}%
Nolan, D., and Temple Lang, D. (2010), ``Computing in the statistics curricula,'' \emph{The American Statistician}, Taylor \& Francis, 64, 97--107.

\leavevmode\hypertarget{ref-pelled2018little}{}%
Pelled, A., Lukito, J., Boehm, F., Yang, J., and Shah, D. (2018), ```Little Marco,'`Lyin'Ted,'`Crooked Hillary,' and the `Biased' media: How Trump used Twitter to attack and organize,'' in \emph{Digital Discussions}, Routledge, pp. 176--196.

\leavevmode\hypertarget{ref-porteous2008fast}{}%
Porteous, I., Newman, D., Ihler, A., Asuncion, A., Smyth, P., and Welling, M. (2008), ``Fast Collapsed Gibbs Sampling for Latent Dirichlet Allocation,'' in \emph{Proceedings of the 14th ACM SIGKDD International Conference on Knowledge Discovery and Data Mining}, pp. 569--577.

\leavevmode\hypertarget{ref-r}{}%
R Core Team (2019), \emph{R: A language and environment for statistical computing}, Vienna, Austria: R Foundation for Statistical Computing.

\leavevmode\hypertarget{ref-drob}{}%
Robinson, D. (2016), ``Text analysis of Trump's tweets confirms he writes only the (angrier) Android half,'' \url{http://varianceexplained.org/r/trump-tweets/}.

\leavevmode\hypertarget{ref-tweet_stream}{}%
``Sampled stream'' (2019), \url{https://developer.twitter.com/en/docs/labs/sampled-stream/overview}.

\leavevmode\hypertarget{ref-tidytext}{}%
Silge, J., and Robinson, D. (2016), ``tidytext: Text mining and analysis using tidy data principles in R,'' \emph{JOSS}, The Open Journal, 1. \url{https://doi.org/10.21105/joss.00037}.

\leavevmode\hypertarget{ref-wells2016trump}{}%
Wells, C., Shah, D. V., Pevehouse, J. C., Yang, J., Pelled, A., Boehm, F., Lukito, J., Ghosh, S., and Schmidt, J. L. (2016), ``How Trump drove coverage to the nomination: Hybrid media campaigning,'' \emph{Political Communication}, Taylor \& Francis, 33, 669--676.

\leavevmode\hypertarget{ref-wickham2019advanced}{}%
Wickham, H. (2019), \emph{Advanced R}, CRC press.

\leavevmode\hypertarget{ref-wickham2016r}{}%
Wickham, H., and Grolemund, G. (2016), \emph{R for Data Science: Import, tidy, transform, visualize, and model data}, O'Reilly Media, Inc.

\leavevmode\hypertarget{ref-wiggins2005understanding}{}%
Wiggins, G., and McTighe, J. (2005), \emph{Understanding by Design}.

\leavevmode\hypertarget{ref-taskscheduleR}{}%
Wijffels, J., and Belmans, O. (2018), \emph{taskscheduleR: Schedule R Scripts and Processes with the Windows Task Scheduler}.

\newpage

\hypertarget{supplementary-files}{%
\section{Supplementary files}\label{supplementary-files}}

\hypertarget{tweets-data-dictionary}{%
\subsection{Tweets data dictionary}\label{tweets-data-dictionary}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \href{https://github.com/fboehm/jse-2019/blob/master/data/tweets-data-dictionary.csv}{Data dictionary}
\end{enumerate}

\hypertarget{r-code-to-reproduce-the-case-study}{%
\subsection{R code to reproduce the case study}\label{r-code-to-reproduce-the-case-study}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \href{https://raw.githubusercontent.com/fboehm/jse-2019/master/Rmd/tweets.Rmd}{tweets.Rmd}
\item
  \href{https://raw.githubusercontent.com/fboehm/jse-2019/master/Rmd/tweets-one.Rmd}{tweets-one.Rmd}
\item
  \href{https://raw.githubusercontent.com/fboehm/jse-2019/master/Rmd/recover_tweets.R}{recover\_tweets.R}
\end{enumerate}

\hypertarget{student-projects}{%
\subsection{Student projects}\label{student-projects}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Student 1 poster: \href{https://github.com/fboehm/jse-2019/blob/master/supplementary/Project_Poster.pdf}{Project\_Poster.pdf}
\item
  Student 1 report: \href{https://github.com/fboehm/jse-2019/blob/master/supplementary/report.pdf}{report.pdf}
\item
  Student 2 useR 2016 slides: \href{https://github.com/fboehm/jse-2019/blob/master/supplementary/user2016boehm.pdf}{user2016boehm.pdf}
\item
  Student 2 poster: \href{https://github.com/fboehm/jse-2019/blob/master/supplementary/warfdiscovery2016boehm.tiff}{warfdiscovery2016boehm.tiff}
\end{enumerate}

\hypertarget{github-repository}{%
\subsection{Github repository}\label{github-repository}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \url{https://github.com/fboehm/jse-2019}
\end{enumerate}

\end{document}