### 2020-11-02 Rmd/reply-to-reviewers-nov2020.Rmd
updated ms AND reply to comments. Updated pdf of ms

---
output: 
  bookdown::pdf_document2:
    toc: false
    number_sections: false
---

November 2, 2020

Dear editors and reviewers:

We are writing to submit an updated manuscript that incorporates the suggested edits.

We include both the text of the comments and our replies.

## Editor's Comments to Author:

* unfortunately, I don't think JSE has formatting guidelines for packages. However, they should be internally consistent in
your paper (e.g., bold, quoted, without format, etc.): rtweet, tidytext, lubridate,...

> We now do this for all packages that we discuss in the text.

* API stands for application programming interface (not application product interface)

> Thank you for this correction. We've made the change.

* how was table 1 ordered? As is, the sorting is (seemingly?) arbitrary, and I think it would be easier to read if sorted. I’d
suggest one of two ways:
- sort by order of category
- sort by the order in which the skills are performed in a standard data analysis.

> Thank you. We now order the rows by category.

* page 10, line 12 “eg., “ should be “e.g.,”

> We've now made this change.

* page 10, line 32. You discovered evidence to “support his strategy” ... what is his strategy? Do you mean to “support his
hypothesis”? or maybe to “support his ideas” ?

> Thank you. We decided to use "support his ideas".

* page 10, line 34 “We share below a case study based on this approach.” I’d clarify to say “Below, we share a case study
on a different widely discussed topic which is analyzed using an approach similar to that from the Super Bowl tweets.”

> Thanks. We've made this change.

* page 12, line 32 ... still something is off as I read the word JSON. You write: “Below is an example of JSON for one
tweet.” Here are some possible alternatives:
- “Below is an example of a JSON entry for one tweet.”
- “Below is an example of a JSON file for one tweet.”
- “Below is an example of a tweet in JSON format.”

> Thank you. We've chosen to use the last of the three suggested wordings.

* “to wrangle the tweet JSON into” (instead change to: “to wrangle the JSON tweet into” or “to wrangle the tweet JSON
data into”) page 13, line 46

> Thanks. We've made this change by using the latter suggested wording.

* I think “document by term matrix” should be “document-term matrix”

> We agree. We've made the changes.

* page 10: suggest changing 'may not be "safe for work"' to "may be "not suitable for work (NSFW)" and add a similar
warning to the Figure 1 caption. Would an additional sentence noting that instructors could have students remove words
from a list of common expletives or inappropriate words?

> Thanks. We've fixed this wording.

* Would it make sense to have all of the supplementary files available on a new public GitHub repo? That would simplify
references and updates.

> Yes, all files are in a Github repo at https://github.com/fboehm/jse-2019. If an additional (new) repo is needed, we're happy to make one.